EDA of DSRCT Samples

### 2024-10-23 analyses/cell-type-dsrct/exploratory_analysis/01-marker-gene-tumor-classification.Rmd
Apply suggestions from code review
modified the saves

Co-authored-by: Joshua Shapiro <josh.shapiro@ccdatalab.org>

```{r}
umap_df_list <- lapply(sce_list, function(x) {
  # get the gene expression counts for all marker genes
  marker_gene_exp <- logcounts(x[marker_genes, ]) |>
    as.matrix() |>
    t() |>
    as.data.frame() |>
    tibble::rownames_to_column("barcodes")

  # pull out the UMAP coordinates and make a data frame to use for plotting
  umap_df <- x |>
    scuttle::makePerCellDF(use.dimred = "UMAP") |>
    # replace UMAP.1 with UMAP1
    dplyr::rename_with(\(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")) |>
    # add in marker gene expression to dataframe
    dplyr::left_join(marker_gene_exp, by = "barcodes") |>
    # combine all genes into a single column for easy faceting
    tidyr::pivot_longer(cols = starts_with("ENSG"),
                        names_to = "ensembl_gene_id",
                        values_to = "gene_expression") |>
    # join with marker gene df to get gene symbols for plotting
    dplyr::left_join(marker_genes_df, by = c("ensembl_gene_id")) |>
    dplyr::select(barcodes,
                  UMAP1,
                  UMAP2,
                  gene_symbol,
                  ensembl_gene_id,
                  gene_expression,
                  cluster)
})

names(umap_df_list) <- metadata_DSRCT$scpca_sample_id
```

@@

  #save plots
  ggsave(
    filename = paste0(
      '../plots/UMAP_marker_expression_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

  #save plots
  ggsave(
    filename = paste0(
      '../plots/UMAP_classification_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

  #save plots
  ggsave(
    filename = paste0('../plots/UMAP_marker_expression_cluster_', names(umap_df_list)[i], '.png'),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

  #save plots
  ggsave(
    filename = paste0(
      '../plots/UMAP_classification_cluster_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )

>>>>

```{r}
umap_df_list <- lapply(sce_list, function(x) {
# pull out the UMAP coordinates and genes and make a data frame to use for plotting
  umap_df <- x |>
    scuttle::makePerCellDF(features = marker_genes, use.dimred = "UMAP") |>
    # replace UMAP.1 with UMAP1
    dplyr::rename_with(\(x) stringr::str_replace(x, "^UMAP\\.", "UMAP")) |>
    # combine all genes into a single column for easy faceting
    tidyr::pivot_longer(cols = starts_with("ENSG"),
                        names_to = "ensembl_gene_id",
                        values_to = "gene_expression") |>
    # join with marker gene df to get gene symbols for plotting
    dplyr::left_join(marker_genes_df, by = c("ensembl_gene_id")) |>
    dplyr::select(barcodes,
                  UMAP1,
                  UMAP2,
                  gene_symbol,
                  ensembl_gene_id,
                  gene_expression,
                  cluster)
})

@@

  #save plots
  ggsave(
    filename = paste0(
      module_base,
      '/plots/UMAP_marker_expression_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

  #save plots
  ggsave(
    filename = paste0(
      module_base,
      '/plots/UMAP_classification_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

  #save plots
  ggsave(
    filename = paste0(module_base, '/plots/UMAP_marker_expression_cluster_', names(umap_df_list)[i], '.png'),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
}

@@

  #save plots
  ggsave(
    filename = paste0(
      module_base,
      '../plots/UMAP_classification_cluster_',
      names(umap_df_list)[i],
      '.png'
    ),
    plot = p,
    width = 6,
    height = 6,
    units = 'in',
    dpi = 150
  )
